---
output:
  github_document:
    toc: yes
    toc_depth: 2
  fig_caption: no
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[![License: GPL v3](https://img.shields.io/badge/License-GPL%20v3-blue.svg)](http://www.gnu.org/licenses/gpl-3.0)

# RSPrePro: Download and preprocess Remote Sensing images (Sentinal-2 and Landsat)

## Warning
This package is under construction: for now only the functions documented in the [function references](http://ggranga.github.io/RSPrePro/reference) are working. 
Moreover, the download and atmospheric correction on Sentinel-2 data [s2tsp](http://github.com/ggranga/s2tsp) was not yet implemented in this R interface, so it must be runned using [python scripts](http://github.com/ggranga/s2tsp).


## Installation

You can install RSPrePro from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("ggranga/RSPrePro")
```

