---
output:
  github_document:
    toc: yes
    toc_depth: 2
  fig_caption: no
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[![Travis-CI Build Status](https://travis-ci.org/ranghetti/sen2r.svg?branch=master)](https://travis-ci.org/ranghetti/sen2r)
[![Codecov coverage status](https://codecov.io/gh/lranghetti/sen2r/branch/master/graph/badge.svg)](https://codecov.io/gh/lranghetti/sen2r)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/sen2r)](https://cran.r-project.org/package=sen2r)
[![License: GPL v3](https://img.shields.io/badge/License-GPL%20v3-blue.svg)](http://www.gnu.org/licenses/gpl-3.0)

# An R toolbox to find, download an preprocess Sentinel-2 data

## Warning
This package is under construction (current version is [pre-release 0.3.1](https://github.com/ranghetti/sen2r/releases/tag/v0.3.1)):
currently only the main processing chain for Sentinel-2 data was implemented.
Please refer to the [crontab of release 0.4.0](https://github.com/ranghetti/sen2r/milestone/3)
to know which implementations will be performed regarding Sentinel-2 data.



## Installation

You can install **sen2r** from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("ranghetti/sen2r")
```

This will install the R package, containing all the functions necessary to preprocess data. 

To download Sentinel-2 images the package makes use of a set of Python functions 
([s2download](https://github.com/ranghetti/s2download)).
These scripts are imported the first time hey are used (alternatively, launch the function
[`s2_download()`](reference/install_s2download.md)
included in the package). 

Atmospheric correction is performed using [sen2cor](http://step.esa.int/main/third-party-plugins-2/sen2cor):
the package will automatically download and install it at first use,
or by running function [`install_sen2cor()`](reference/install_sen2cor.md).
Please notice that the use of sen2cor algorythm was not yet possible under MAC.

Preprocessing functions make use of [GDAL](http://www.gdal.org), which must 
support JPEG2000 format. On Windows, it is strongly recommended to install it 
using the [OSGeo4W installer](http://download.osgeo.org/osgeo4w/osgeo4w-setup-x86_64.exe)
in advanced mode, and checking the installation of `openjpeg` library.

## Usage

The simpler way to use **sen2r** is to execute the function `sen2r()` 
without any argument: this opens the GUI to select the processing parameters,
and then launches the main function.
    
Alternatively, 
[`sen2r()`](reference/sen2r.md)
can be launched with a list of parameters (created with 
[`s2_gui()`](reference/s2_gui.md))
or passing manually the parameters as arguments of the function
(see the documentation of the function for further details).
    
Other specific functions can be used to run single steps separately:

* [`s2_list()`](reference/s2_list.md)
    to retrieve the list of available Sentinel-2 products basing on input parameters;
* [`s2_download()`](reference/s2_download.md)
    to download Sentinel-2 products;
* [`sen2cor()`](reference/sen2cor.html)
    to correct level-1C products using [sen2cor](http://step.esa.int/main/third-party-plugins-2/sen2cor);
* [`s2_translate()`](reference/s2_translate.md)
    to convert Sentinel-2 products from SAFE format to a format managed by GDAL;
* [`s2_merge()`](reference/s2_merge.md)
    to merge Sentinel-2 tiles which have the same date and orbit;
* [`gdal_warp()`](reference/gdal_warp.md)
    to clip, reproject and warp raster files (this is a wrapper to call
    [gdal_translate](http://www.gdal.org/gdal_translate.html) or 
    [gdalwarp](http://www.gdal.org/gdalwarp.html) basing on input parameters);
* [`s2_mask()`](reference/s2_mask.md)
    to apply a cloud mask to Sentinel-2 products;
* [`s2_calcindices()`](reference/s2_calcindices.md)
    to compute maps of spectral indices from Sentinel-2 Surface Reflectance multiband raster files.

## Credits

**sen2r** is being developed by Luigi Ranghetti and Lorenzo Busetto
([IREA-CNR](http://www.irea.cnr.it)),
and it is released under [GPL 3.0](https://www.gnu.org/licenses/gpl.html).


