---
title: "Installation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installation}
  %\VignetteEngine{knitr::knitr}
  %\usepackage[UTF-8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Instruction are provided separately for each OS:

- [Install on Linux systems](#linux)
- [Install on Windows](#windows)

The support for MacOS will be added in future.

Alternatively, <span style="color:#5793dd;vertical-align:top;font-size:90%;font-weight:normal;">sen</span><span style="color:#6a7077;vertical-align:baseline;font-size:115%;font-weight:bolder;">2</span><span style="color:#2f66d5;vertical-align:baseline;font-size:90%;font-weight:bold;">r</span> can be used [in a Docker container](docker.md).


## <a id="linux"></a>On Linux systems

### Installing dependencies

Before installing <span style="color:#5793dd;vertical-align:top;font-size:90%;font-weight:normal;">sen</span><span style="color:#6a7077;vertical-align:baseline;font-size:115%;font-weight:bolder;">2</span><span style="color:#2f66d5;vertical-align:baseline;font-size:90%;font-weight:bold;">r</span>, it is recommended to check [R](https://cran.r-project.org/bin/linux) (v. 3.4.4 or higher) and the following libraries to be present:

- [GDAL](https://www.gdal.org), [PROJ](https://proj4.org) and [GEOS](https://trac.osgeo.org/geos)
    (required by [rgdal](https://r-forge.r-project.org/projects/rgdal)
    and [rgeos](https://r-forge.r-project.org/projects/rgeos));
- [jq](https://stedolan.github.io/jq)
    (required by [jqr](https://github.com/ropensci/jqr));
- [protobuf](https://github.com/protocolbuffers/protobuf)
    (required by [protolite](https://github.com/jeroen/protolite));
- [V8](https://v8.dev) 3.14 branch
    (required by [geojsonio](https://github.com/ropensci/geojsonio));
- [openssl](https://www.openssl.org)
    (required by [openssl](https://github.com/jeroen/openssl));
- [cairo](https://www.cairographics.org)
    (required by [gdtools](https://github.com/davidgohel/gdtools));
- [libcurl](https://curl.haxx.se/libcurl)
    (required by [curl](https://github.com/jeroen/curl));
- [NetCDF](https://www.unidata.ucar.edu/software/netcdf)
    (required by [ncdf4](https://github.com/mdsumner/ncdf4));
- [libxml2](http://xmlsoft.org)
    (required by [XML](https://cran.r-project.org/web/packages/XML/index.html)).
    
Other binaries are required (or suggested), since they are used in runtime by the package:

* [**GDAL**](http://www.gdal.org) (with support for JP2OpenJPEG format): 
    this is a mandatory dependency, needed for all the processing operations 
    and to retrieve metadata from SAFE products;
* [**sen2cor**](http://step.esa.int/main/third-party-plugins-2/sen2cor) 
    is used to perform atmospheric correction of Sentinel-2 Level-1C products: 
    it is required by the package, unless you choose not to correct products locally 
    (using only Level-1C – TOA products or dowloading directly Level-2A products).
* [**aria2**](https://aria2.github.io) is an alternative downloader which
    can be used to faster the download of SAFE archives; it can be optionally
    installed and used.

To install both package dependencies and runtime binaries (excepting **sen2cor**, which is installed by <span style="color:#5793dd;vertical-align:top;font-size:90%;font-weight:normal;">sen</span><span style="color:#6a7077;vertical-align:baseline;font-size:115%;font-weight:bolder;">2</span><span style="color:#2f66d5;vertical-align:baseline;font-size:90%;font-weight:bold;">r</span>):

- On Debian and Ubuntu:
    ```{bash, eval = FALSE}
    sudo apt-get install -y r-base gdal-bin python-gdal aria2 libpython-dev libudunits2-dev libgdal-dev libjq-dev libprotobuf-dev protobuf-compiler libv8-3.14-dev libssl-dev libcairo2-dev
    ```
    On Debian, `proj-epsg` is also required:
    ```{bash, eval = FALSE}
    sudo apt-get install -y proj-epsg
    ```
    On Ubuntu 16.04 LTS Xenial and 14.04 LTS Trusty, the repositories [ubuntugis](https://launchpad.net/~ubuntugis) and [opencpu](https://launchpad.net/~opencpu) must be added before installing the packages:
    ```{bash, eval = FALSE}
    sudo add-apt-repository -y ppa:opencpu/jq
    sudo add-apt-repository -y ppa:ubuntugis/ubuntugis-unstable
    sudo apt-get update -q
    ```
    Installation was tested over Debian 9 Stretch and Ubuntu 18.04 LTS Bionic; newer versions (Ubuntu > 18.04, Debian > 9) should work, like Ubuntu 16.04 LTS Xenial and 14.04 LTS Trusty should also do. Other releases (Debian &#8804; 8; Ubuntu non-LTS < 18.04; Ubuntu LTS &#8804; 12.04) are not supported.
- On Fedora:
    ```{bash, eval = FALSE}
    sudo yum install R gdal gdal-python aria2 python-devels gdal-devel proj-devel geos-devel udunits2-devel jq-devel protobuf-devel v8-314-devel openssl-devel cairo-devel libcurl-devel netcdf-devel libxml2-devel
    ```
    Installation was tested over Fedora 29.
- On ArchLinux:
    ```{bash, eval = FALSE}
    sudo pacman -S --noconfirm r base-devel gcc-fortran python-gdal wget aria2 jq protobuf git
    cd /tmp; git clone https://aur.archlinux.org/gyp-git.git
    cd gyp-git; makepkg -si --noconfirm
    cd /tmp; git clone https://aur.archlinux.org/v8-3.14.git
    cd v8-3.14; makepkg -si --noconfirm
    cd /tmp; git clone https://aur.archlinux.org/udunits.git
    cd udunits; makepkg -si --noconfirm
    ```
    Installation was tested on date 2018-11-21.

### Installing <span style="color:#5793dd;vertical-align:top;font-size:90%;font-weight:normal;">sen</span><span style="color:#6a7077;vertical-align:baseline;font-size:115%;font-weight:bolder;">2</span><span style="color:#2f66d5;vertical-align:baseline;font-size:90%;font-weight:bold;">r</span>

Before proceeding, it is suggested to update the existing R packages:
```{r, eval = FALSE}
update.packages(ask = FALSE)
```
Doing that, the following warning could appear:
```
Warning in install.packages(update[instlib == l, "Package"], l, contriburl = contriburl,  :
  'lib = "/usr/lib/R/library"' is not writable
Would you like to use a personal library instead?  (y/n)
```
If so, press `y` and continue.
Moreover, if no personal libraries already exist, it is asked to create a new one: ask `y` and continue.

After that, the package can be installed from GitHub with the R package **remotes**.
To do it:

1. install the package **remotes**, if missing:
    ```{r, eval = FALSE}
    install.packages("remotes")
    ```
2. load it and install 
    <span style="color:#5793dd;vertical-align:top;font-size:90%;font-weight:normal;">sen</span><span
    style="color:#6a7077;vertical-align:baseline;font-size:115%;font-weight:bolder;">2</span><span
    style="color:#2f66d5;vertical-align:baseline;font-size:90%;font-weight:bold;">r</span>:
    ```{r, eval = FALSE}
    library(remotes)
    install_github("ranghetti/sen2r")
    ```

This will install the R package along with its package dependencies,
containing all the functions necessary to preprocess data. 

Finally, **sen2cor** can be installed from R:
```{r, eval = FALSE}
library(sen2r)
install_sen2cor()
```

## <a id="windows"></a>On Windows

### Installing <span style="color:#5793dd;vertical-align:top;font-size:90%;font-weight:normal;">sen</span><span style="color:#6a7077;vertical-align:baseline;font-size:115%;font-weight:bolder;">2</span><span style="color:#2f66d5;vertical-align:baseline;font-size:90%;font-weight:bold;">r</span>

[R](https://cran.r-project.org/bin/windows/base) must be installed before proceeding.

In the case it was already installed, it is suggested to update the existing R packages:
```{r, eval = FALSE}
update.packages(ask = FALSE)
```

<span style="color:#5793dd;vertical-align:top;font-size:90%;font-weight:normal;">sen</span><span style="color:#6a7077;vertical-align:baseline;font-size:115%;font-weight:bolder;">2</span><span style="color:#2f66d5;vertical-align:baseline;font-size:90%;font-weight:bold;">r</span> can be installed from GitHub with the R package **remotes**.
To do it:

1. install the package **remotes**, if missing:
    ```{r, eval = FALSE}
    install.packages("remotes")
    ```
2. load it and install 
    <span style="color:#5793dd;vertical-align:top;font-size:90%;font-weight:normal;">sen</span><span
    style="color:#6a7077;vertical-align:baseline;font-size:115%;font-weight:bolder;">2</span><span
    style="color:#2f66d5;vertical-align:baseline;font-size:90%;font-weight:bold;">r</span>:
    ```{r, eval = FALSE}
    library(remotes)
    install_github("ranghetti/sen2r")
    ```

This will install the R package along with its package dependencies,
containing all the functions necessary to preprocess data. 

### Installing runtime dependencies

To run the functions correctly, some external dependences are required:

* [**GDAL**](http://www.gdal.org) (with support for JP2OpenJPEG format): 
    this is a mandatory dependency, needed for all the processing operations 
    and to retrieve metadata from SAFE products;
* [**sen2cor**](http://step.esa.int/main/third-party-plugins-2/sen2cor) 
    is used to perform atmospheric correction of Sentinel-2 Level-1C products: 
    it is required by the package, unless you choose not to correct products locally 
    (using only Level-1C – TOA products or dowloading directly Level-2A products).
* **Wget** is the downloader used by the package; it is required to work online.
* [**aria2**](https://aria2.github.io) is an alternative downloader which
    can be used to faster the download of SAFE archives; it can be optionally
    installed and used.

These dependences can be graphically checked launching the function
```{r, eval = FALSE}
library(sen2r)
check_sen2r_deps()
```
This function opens a GUI which help to check that they are satisfied;
if some of them are missing, the GUI allows to install them directly.

<!-- Atmospheric correction is performed using [sen2cor](http://step.esa.int/main/third-party-plugins-2/sen2cor): 
the package will automatically download and install it at first use,
or by running function [`install_sen2cor()`](reference/install_sen2cor.md).
Please notice that the use of sen2cor algorythm was not yet possible under MAC.

Preprocessing functions make use of [GDAL](http://www.gdal.org), which must 
support JPEG2000 format. On Windows, it is strongly recommended to install it 
using the [OSGeo4W installer](http://download.osgeo.org/osgeo4w/osgeo4w-setup-x86_64.exe)
in advanced mode, and checking the installation of `openjpeg` library. -->

## <a id="mac"></a>On MacOS

The support for MacOS will be added in future.