---
title: "Output file structure"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using sen2r() from the command line}
  %\VignetteEngine{knitr::knitr}
  %\usepackage[UTF-8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Products generated using 
<span style="color:#5793dd;vertical-align:top;font-size:90%;font-weight:normal;">sen</span><span style="color:#6a7077;vertical-align:baseline;font-size:115%;font-weight:bolder;">2</span><span style="color:#2f66d5;vertical-align:baseline;font-size:90%;font-weight:bold;">r</span>
functions follows a precise [naming convention](#naming-convention),
and are organised within a customisable [folder structure](#folder-structure).


## Naming convention

ESA Sentinel-2 naming convention is particularly long-winded; so, a different
convention was adopted, follows this schema:

<span style='font-family:monospace;font-weight:bold;'><span style='color:#FF3300;'>S2mll</span>\_<span style='color:#FF3300;'>yyyymmdd</span>\_<span style='color:#FF3300;'>rrr</span>\_<span style='color:#FF3300;'>EXTNAME</span>\_<span style='color:#FF3300;'>PROD</span>\_<span style='color:#FF3300;'>rr</span>\.<span style='color:#FF3300;'>fff</span></span>

(e.g. <span style='font-family:monospace;'><span style='color:#FF6611;'>S2A2A</span>\_<span style='color:#FF6611;'>20190723</span>\_<span style='color:#FF6611;'>022</span>\_<span style='color:#FF6611;'>Barbellino</span>\_<span style='color:#FF6611;'>BOA</span>\_<span style='color:#FF6611;'>10</span>\.<span style='color:#FF6611;'>tif</span></span>)

 * <span style='color:#FF3300;font-family:monospace;font-weight:bold;'>S2mll</span>
     (length: 5) shows the mission ID 
     (<span style='font-family:monospace;color:#FF6611;'>S2A</span> or 
     <span style='font-family:monospace;color:#FF6611;'>S2B</span>) 
     and the product level 
     (<span style='font-family:monospace;color:#FF6611;'>1C</span> or 
     <span style='font-family:monospace;color:#FF6611;'>2A</span>);
 * <span style='color:#FF3300;font-family:monospace;font-weight:bold;'>yyyymmdd</span>
     (length: 8) is the sensing date (e.g. 
     <span style='font-family:monospace;color:#FF6611;'>20190723</span> for 2019-07-23); 
     the hour is skipped, since a single sensor can not pass 
     two times in a day on the same tile);
 * <span style='color:#FF3300;font-family:monospace;font-weight:bold;'>rrr</span>
     (length: 3) is the relative orbit number (e.g. 
     <span style='font-family:monospace;color:#FF6611;'>022</span>);
 * <span style='color:#FF3300;font-family:monospace;font-weight:bold;'>EXTNAME</span>
     (length: variable, see [details](#extent-name))
     is the output extent name (e.g. 
     <span style='font-family:monospace;color:#FF6611;'>Barbellino</span>);
     it could also be the the tile number (length: 5 -- e.g. 
     <span style='font-family:monospace;color:#FF6611;'>32TQQ</span>)
     for outputs of function `s2_translate()`,
     and is kept empty for outputs of function `s2_merge()`;
 * <span style='color:#FF3300;font-family:monospace;font-weight:bold;'>PROD</span>
     (length: variable) is the output product name, being one of these:
     - an output product: 
         <span style='font-family:monospace;color:#FF6611;'>TOA</span>
         (13-bands Top-Of-Atmosphere Reflectance),
         <span style='font-family:monospace;color:#FF6611;'>BOA</span>
         (13-bands Bottom-Of-Atmosphere Reflectance) or 
         <span style='font-family:monospace;color:#FF6611;'>SCL</span>
         (Scene Classification Map);
     - a spectral index name (the list of supported index names can be obtained
         with the function `list_indices("name")`);
     - the name of a RGB image (length: 7, see [details](#rgb-names)).
 * <span style='color:#FF3300;font-family:monospace;font-weight:bold;'>rr</span>
     (length: 2) is the original minimum spatial resolution in metres
     (<span style='font-family:monospace;color:#FF6611;'>10</span>, 
     <span style='font-family:monospace;color:#FF6611;'>20</span> or 
     <span style='font-family:monospace;color:#FF6611;'>60</span>);
 * <span style='color:#FF3300;font-family:monospace;font-weight:bold;'>fff</span>
     (length: variable, generally 3) is the file extension
     (e.g. <span style='font-family:monospace;color:#FF6611;'>tif</span>).


### Extent name
The extent name is defined with the argument `extent_name` of function `sen2r()`.

The user can choose any non-empty alphanumeric string as extent name, as long 
as it does not contain spaces, points nor underscores, and as long as it is not 
a five-length string with the same structure of a tile ID
(two numeric and three uppercase character values), 
because this structure is reserved for intermediate layers.


### RGB names
The names of RGB images are 7-length strings following the structure
<span style='font-family:monospace;font-weight:bold;color:#FF3300;'>RGBrgbX</span>
(e.g. <span style='font-family:monospace;color:#FF6611;'>RGBb84B</span>),
where:

- <span style='color:#FF3300;font-family:monospace;font-weight:bold;'>r</span>, 
    <span style='color:#FF3300;font-family:monospace;font-weight:bold;'>g</span> and
    <span style='color:#FF3300;font-family:monospace;font-weight:bold;'>b</span>
    are the number of the bands to be used respectively for red, green and blue, 
    in hexadecimal format (in the example above, 
    <span style='font-family:monospace;color:#FF6611;'>b</span>, 
    <span style='font-family:monospace;color:#FF6611;'>8</span> and 
    <span style='font-family:monospace;color:#FF6611;'>4</span>
    correspond to bands 11, 8 and 4);
- <span style='color:#FF3300;font-family:monospace;font-weight:bold;'>X</span>
    is <span style='font-family:monospace;color:#FF6611;'>B</span> if index source is BOA, 
    or <span style='font-family:monospace;color:#FF6611;'>T</span> if source is TOA.


## Folder structure

Processing functions allows setting an output directory using the arguments 
`path_out`, `path_rgb` and `path_indices` (`sen2r()`) or `outdir` 
(`s2_translate()`, `s2_merge()`, `s2_mask()`, `s2_rgb()` and `s2_calcindices()`).
These folders are automatically created if their parent directories exist.

By default, functions create a subfolder for each product / index / RGB type,
in which files are stored.
Each product subfolder includes a sub-subfolder named `thumbnails`,
which contains small JPEG or PNG images which can be used to rapidly monitor
what was done (see `s2_thumbnails()` for further details).

For example, the following command would produce the outputs listed below.
```{r, eval=FALSE}
json_path <- build_example_param_file()
out_dir_1 <- tempfile(pattern = "sen2r_out_1_")

library(sen2r)
sen2r(
  json_path, 
  timewindow = c("2019-07-13","2019-07-23"),
  path_out = out_dir_1
)
```
```
[2020-02-06 09:42:40] #### Starting sen2r execution. ####
[2020-02-06 09:42:40] Searching for available SAFE products on SciHub...
[2020-02-06 09:42:46] Computing output names...
[2020-02-06 09:42:48] Processing group 1 of 5...
```
_Most of output message were skipped;_
_see [this vignette](sen2r_cmd.html) for further details._
```
╔══════════════════════════════════════════════════════════════════════════════
║ sen2r Processing Report
╟──────────────────────────────────────────────────────────────────────────────
║ Dates to be processed based on processing parameters: 5
║ Processing completed for: all expected dates.
╚══════════════════════════════════════════════════════════════════════════════
[2020-02-06 09:44:45] #### sen2r session terminated. ####
The processing chain can be re-launched with the command:
  sen2r("/home/lranghetti/.sen2r/proc_par/s2proc_20200206_094240.json")
```
```{r, eval=FALSE}
system(paste("tree", out_dir_1)) # working on Linux with binary "tree" installed
```
```
/tmp/RtmpUlAdHw/sen2r_out_1_a55119f4ed09
├── BOA
│   ├── S2A2A_20190713_022_sen2r_BOA_10.tif
│   ├── S2A2A_20190716_065_sen2r_BOA_10.tif
│   ├── S2A2A_20190723_022_sen2r_BOA_10.tif
│   ├── S2B2A_20190718_022_sen2r_BOA_10.tif
│   ├── S2B2A_20190721_065_sen2r_BOA_10.tif
│   └── thumbnails
│       ├── S2A2A_20190713_022_sen2r_BOA_10.jpg
│       ├── S2A2A_20190713_022_sen2r_BOA_10.jpg.aux.xml
│       ├── S2A2A_20190716_065_sen2r_BOA_10.jpg
│       ├── S2A2A_20190716_065_sen2r_BOA_10.jpg.aux.xml
│       ├── S2A2A_20190723_022_sen2r_BOA_10.jpg
│       ├── S2A2A_20190723_022_sen2r_BOA_10.jpg.aux.xml
│       ├── S2B2A_20190718_022_sen2r_BOA_10.jpg
│       ├── S2B2A_20190718_022_sen2r_BOA_10.jpg.aux.xml
│       ├── S2B2A_20190721_065_sen2r_BOA_10.jpg
│       └── S2B2A_20190721_065_sen2r_BOA_10.jpg.aux.xml
├── MSAVI2
│   ├── S2A2A_20190713_022_sen2r_MSAVI2_10.tif
│   ├── S2A2A_20190716_065_sen2r_MSAVI2_10.tif
│   ├── S2A2A_20190723_022_sen2r_MSAVI2_10.tif
│   ├── S2B2A_20190718_022_sen2r_MSAVI2_10.tif
│   ├── S2B2A_20190721_065_sen2r_MSAVI2_10.tif
│   └── thumbnails
│       ├── S2A2A_20190713_022_sen2r_MSAVI2_10.jpg
│       ├── S2A2A_20190713_022_sen2r_MSAVI2_10.jpg.aux.xml
│       ├── S2A2A_20190716_065_sen2r_MSAVI2_10.jpg
│       ├── S2A2A_20190716_065_sen2r_MSAVI2_10.jpg.aux.xml
│       ├── S2A2A_20190723_022_sen2r_MSAVI2_10.jpg
│       ├── S2A2A_20190723_022_sen2r_MSAVI2_10.jpg.aux.xml
│       ├── S2B2A_20190718_022_sen2r_MSAVI2_10.jpg
│       ├── S2B2A_20190718_022_sen2r_MSAVI2_10.jpg.aux.xml
│       ├── S2B2A_20190721_065_sen2r_MSAVI2_10.jpg
│       └── S2B2A_20190721_065_sen2r_MSAVI2_10.jpg.aux.xml
├── NDVI
│   ├── S2A2A_20190713_022_sen2r_NDVI_10.tif
│   ├── S2A2A_20190716_065_sen2r_NDVI_10.tif
│   ├── S2A2A_20190723_022_sen2r_NDVI_10.tif
│   ├── S2B2A_20190718_022_sen2r_NDVI_10.tif
│   ├── S2B2A_20190721_065_sen2r_NDVI_10.tif
│   └── thumbnails
│       ├── S2A2A_20190713_022_sen2r_NDVI_10.jpg
│       ├── S2A2A_20190713_022_sen2r_NDVI_10.jpg.aux.xml
│       ├── S2A2A_20190716_065_sen2r_NDVI_10.jpg
│       ├── S2A2A_20190716_065_sen2r_NDVI_10.jpg.aux.xml
│       ├── S2A2A_20190723_022_sen2r_NDVI_10.jpg
│       ├── S2A2A_20190723_022_sen2r_NDVI_10.jpg.aux.xml
│       ├── S2B2A_20190718_022_sen2r_NDVI_10.jpg
│       ├── S2B2A_20190718_022_sen2r_NDVI_10.jpg.aux.xml
│       ├── S2B2A_20190721_065_sen2r_NDVI_10.jpg
│       └── S2B2A_20190721_065_sen2r_NDVI_10.jpg.aux.xml
├── RGB432B
│   ├── S2A2A_20190713_022_sen2r_RGB432B_10.tif
│   ├── S2A2A_20190716_065_sen2r_RGB432B_10.tif
│   ├── S2A2A_20190723_022_sen2r_RGB432B_10.tif
│   ├── S2B2A_20190718_022_sen2r_RGB432B_10.tif
│   ├── S2B2A_20190721_065_sen2r_RGB432B_10.tif
│   └── thumbnails
│       ├── S2A2A_20190713_022_sen2r_RGB432B_10.jpg
│       ├── S2A2A_20190713_022_sen2r_RGB432B_10.jpg.aux.xml
│       ├── S2A2A_20190716_065_sen2r_RGB432B_10.jpg
│       ├── S2A2A_20190716_065_sen2r_RGB432B_10.jpg.aux.xml
│       ├── S2A2A_20190723_022_sen2r_RGB432B_10.jpg
│       ├── S2A2A_20190723_022_sen2r_RGB432B_10.jpg.aux.xml
│       ├── S2B2A_20190718_022_sen2r_RGB432B_10.jpg
│       ├── S2B2A_20190718_022_sen2r_RGB432B_10.jpg.aux.xml
│       ├── S2B2A_20190721_065_sen2r_RGB432B_10.jpg
│       └── S2B2A_20190721_065_sen2r_RGB432B_10.jpg.aux.xml
├── RGB843B
│   ├── S2A2A_20190713_022_sen2r_RGB843B_10.tif
│   ├── S2A2A_20190716_065_sen2r_RGB843B_10.tif
│   ├── S2A2A_20190723_022_sen2r_RGB843B_10.tif
│   ├── S2B2A_20190718_022_sen2r_RGB843B_10.tif
│   ├── S2B2A_20190721_065_sen2r_RGB843B_10.tif
│   └── thumbnails
│       ├── S2A2A_20190713_022_sen2r_RGB843B_10.jpg
│       ├── S2A2A_20190713_022_sen2r_RGB843B_10.jpg.aux.xml
│       ├── S2A2A_20190716_065_sen2r_RGB843B_10.jpg
│       ├── S2A2A_20190716_065_sen2r_RGB843B_10.jpg.aux.xml
│       ├── S2A2A_20190723_022_sen2r_RGB843B_10.jpg
│       ├── S2A2A_20190723_022_sen2r_RGB843B_10.jpg.aux.xml
│       ├── S2B2A_20190718_022_sen2r_RGB843B_10.jpg
│       ├── S2B2A_20190718_022_sen2r_RGB843B_10.jpg.aux.xml
│       ├── S2B2A_20190721_065_sen2r_RGB843B_10.jpg
│       └── S2B2A_20190721_065_sen2r_RGB843B_10.jpg.aux.xml
└── SCL
    ├── S2A2A_20190713_022_sen2r_SCL_10.tif
    ├── S2A2A_20190716_065_sen2r_SCL_10.tif
    ├── S2A2A_20190723_022_sen2r_SCL_10.tif
    ├── S2B2A_20190718_022_sen2r_SCL_10.tif
    ├── S2B2A_20190721_065_sen2r_SCL_10.tif
    └── thumbnails
        ├── S2A2A_20190713_022_sen2r_SCL_10.png
        ├── S2A2A_20190713_022_sen2r_SCL_10.png.aux.xml
        ├── S2A2A_20190716_065_sen2r_SCL_10.png
        ├── S2A2A_20190716_065_sen2r_SCL_10.png.aux.xml
        ├── S2A2A_20190723_022_sen2r_SCL_10.png
        ├── S2A2A_20190723_022_sen2r_SCL_10.png.aux.xml
        ├── S2B2A_20190718_022_sen2r_SCL_10.png
        ├── S2B2A_20190718_022_sen2r_SCL_10.png.aux.xml
        ├── S2B2A_20190721_065_sen2r_SCL_10.png
        └── S2B2A_20190721_065_sen2r_SCL_10.png.aux.xml

12 directories, 90 files
```
In each of the 6 subfolders `BOA`, `MSAVI2`, `NDVI`, `RGB432B`, `RGB843B` and `SCL`
5 products (one for each date) were created.
Sub-subfolders `thumbnails` contains thumbnail images.

Users can skip the creation of products' subfolders setting the argument
`sen2r(..., path_subdirs = FALSE)` (default: `TRUE`), or doing the same with
the argument `subdir` of other processing functions.
The creation of thumbnails can be skipped using `sen2r(..., thumbnails = FALSE)`
(default: `TRUE`).
Both the options can be also set [using the `sen2r()` GUI](sen2r_gui.html).

For example, the following command would produce the outputs listed below.
```{r, eval=FALSE}
out_dir_2 <- tempfile(pattern = "sen2r_out_2_")

sen2r(
  json_path, 
  timewindow = c("2019-07-13","2019-07-23"),
  path_out = out_dir_2,
  path_subdirs = FALSE,
  thumbnails = FALSE
)
```
```
[2020-02-06 09:52:30] #### Starting sen2r execution. ####
[2020-02-06 09:52:30] Searching for available SAFE products on SciHub...
[2020-02-06 09:52:35] Computing output names...
[2020-02-06 09:52:36] Processing group 1 of 5...

# most of output message were skipped

╔══════════════════════════════════════════════════════════════════════════════
║ sen2r Processing Report
╟──────────────────────────────────────────────────────────────────────────────
║ Dates to be processed based on processing parameters: 5
║ Processing completed for: all expected dates.
╚══════════════════════════════════════════════════════════════════════════════
[2020-02-06 09:54:16] #### sen2r session terminated. ####
The processing chain can be re-launched with the command:
  sen2r("/home/lranghetti/.sen2r/proc_par/s2proc_20200206_095230.json")
```
```{r, eval=FALSE}
system(paste("tree", out_dir_2))
```
```
/tmp/RtmpUlAdHw/sen2r_out_2_a5513ad348b0
├── S2A2A_20190713_022_sen2r_BOA_10.tif
├── S2A2A_20190713_022_sen2r_MSAVI2_10.tif
├── S2A2A_20190713_022_sen2r_NDVI_10.tif
├── S2A2A_20190713_022_sen2r_RGB432B_10.tif
├── S2A2A_20190713_022_sen2r_RGB843B_10.tif
├── S2A2A_20190713_022_sen2r_SCL_10.tif
├── S2A2A_20190716_065_sen2r_BOA_10.tif
├── S2A2A_20190716_065_sen2r_MSAVI2_10.tif
├── S2A2A_20190716_065_sen2r_NDVI_10.tif
├── S2A2A_20190716_065_sen2r_RGB432B_10.tif
├── S2A2A_20190716_065_sen2r_RGB843B_10.tif
├── S2A2A_20190716_065_sen2r_SCL_10.tif
├── S2A2A_20190723_022_sen2r_BOA_10.tif
├── S2A2A_20190723_022_sen2r_MSAVI2_10.tif
├── S2A2A_20190723_022_sen2r_NDVI_10.tif
├── S2A2A_20190723_022_sen2r_RGB432B_10.tif
├── S2A2A_20190723_022_sen2r_RGB843B_10.tif
├── S2A2A_20190723_022_sen2r_SCL_10.tif
├── S2B2A_20190718_022_sen2r_BOA_10.tif
├── S2B2A_20190718_022_sen2r_MSAVI2_10.tif
├── S2B2A_20190718_022_sen2r_NDVI_10.tif
├── S2B2A_20190718_022_sen2r_RGB432B_10.tif
├── S2B2A_20190718_022_sen2r_RGB843B_10.tif
├── S2B2A_20190718_022_sen2r_SCL_10.tif
├── S2B2A_20190721_065_sen2r_BOA_10.tif
├── S2B2A_20190721_065_sen2r_MSAVI2_10.tif
├── S2B2A_20190721_065_sen2r_NDVI_10.tif
├── S2B2A_20190721_065_sen2r_RGB432B_10.tif
├── S2B2A_20190721_065_sen2r_RGB843B_10.tif
└── S2B2A_20190721_065_sen2r_SCL_10.tif

0 directories, 30 files
```
